---
title: "Project 1"
author: "Gabriela Levenfeld Sabau"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: false
    number_sections: false
    highlight: tango
    latex_engine: xelatex
    fig_caption: true
    extra_dependencies:
      - geometry
bibliography: references.bib
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r load_libraries, include=FALSE}
library(skimr)
library(caret) # For split the dataset
library(tidyverse) # For glimpse
```

# Telco Customer Churn

### Available Data

TODO: Write
- Link source
- Pequeña descripción sobre el dataset

## Download data

```{r load_data}
data <- read.csv("Telco-Customer-Churn.csv", header = TRUE,sep = ",")
```


```{r small_EDA}
# Small Exploratory Data Analysis
skim(data)
# Presents results for every column; the statistics it provides depend on the class of the variable

glimpse(data)

# For visualize the first 2 rows
head(data, 2)
```


```{r}
# Counting NA
colSums(is.na(data)) # 11 NA's in TotalCharges

colSums(is.na(data))/nrow(data)
percentage_na <- colSums(is.na(data))/nrow(data)*100 # In %
sort(percentage_na, decreasing = TRUE)
```

```{r}
table(data$Churn)
```

```{r}
# Split into training and testing set
set.seed(1234) # For reproducibility
index <- createDataPartition(data$Churn, p=0.8, list=FALSE)

X_train <- data[index,]
X_test <- data[-index,]
y_train <- data[index,]$Churn
y_test <- data[-index,]$Churn

# In order to check proportion are still the same
table(data$Churn)/length(data$Churn)
table(X_train$Churn)/length(X_train$Churn)
table(X_test$Churn)/length(X_test$Churn)
```


### Exploratory Data Analysis (EDA)

*Some ideas*
REVIEW: some ideas to check for
Working with real data can be challenging-
The data can be:
- Be stored in different formats.
- Being incomplete: missing values.
- Noise (data contaminated): may have errors.
- Inconsistend: not corresponding to the attribute domain or contradictory

# Statistical tools



## References

```{r}
# Returns the names of all packages loaded in the current R session
knitr::write_bib(.packages(), "references.bib")
```

<div id="refs"></div>
